Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-06-30 Manuscript/MPESubmission/Revision/response.tex

\item{\textcolor{blue}{\textbf{line 595-606:} this paragraph seems unsatisfactory advice for practitioners. First, advising that the tree topology be fixed using the Bayesian consensus tree before conducting analyses such as dating strongly conditions the results on only a single tree and as the authors point out in the parenthetical statement, the dating information could improve accuracy. Indeed one of the many advantages of tip-dating is the joint inference of tree topology and divergence times. The following statements that the posterior distribution should not be discarded does not leave the reader with a clear path that they should follow - if only the fixed tree topology is used to estimate divergence times, how can the substitution rate-scaled posterior distribution of highly probable trees be used for most comparative analyses which require time-scaled branch lengths? This paragraph requires further justification and this study does have important bearing on a subject that gets little attention. }}

%To do (rewrite)

\item{\textcolor{blue}{\textbf{line 612-613:} too obvious and not especially helpful. We should aim to have as little missing data as possible…more helpful, and something especially discovered in this study, is concentrating on having overlapping morphological characters between the living and fossil taxa to be able to place the living taxa with strong support via large datasets (combined data) and being able to position the fossils based on their shared derived characters with living taxa. This is an important point to make clearer. }}

\item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}

We modified the sentence as follow: "Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa." (line @@@),

\item{\textcolor{blue}{\textbf{line 615-618:} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max. likelihood, other than to re-hash that the authors feel the Bayesian consensus should be used to fix topologies for other downstream analyses. For example, can the authors cite some combined analyses that used only likelihood analyses and came up with controversial results? Or what about other studies that have done combined analyses with different optimality criteria and found different topologies? What about implications for studies that have used molecular scaffold techniques to fix the topology of living taxa based on analyses of only DNA analyses and then fit the fossils into this fixed tree? Also 'further' is misspelled.  }}

%To do (rewrite) - typo ok

\item{\textcolor{blue}{\textbf{line 628:} use of computer cluster may be better placed in the methods, depending on the norm for MPE.}}

The Lonsdale cluster requires this specific line in the acknowledgements of any study using it's infrastructure.

\item{\textcolor{blue}{\textbf{Figure 3:} Do not repeat the aside to similarity with t-test again from Fig 2 and text.}}

We removed the sentence about the t-test similarity from both figures captions. %TG: not sure if that's what the reviewer wanted.

\item{\textcolor{blue}{\textbf{Figure 4:} color codes in legend do not match up to the colors in the PDF I received. I don't know if this is an error in conversion, but the colors are black, red, green and blue in my PDF. The legend states the following: 'Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue)' which doesn't seem correct since blue is repeated. This makes interpreting the figure difficult for me at this time, but I assume given the text in the Results that the Bayesian methods are black and red in the figure I received, indicating they were closer to the best tree than max likelihood trees. Red and green should not be used together.}}

We apologies for this silly mistake and thanks the reviewer for pointing it out. We changed the color scheme to "Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange)".
%To do: change the graphs colours

\item{\textcolor{blue}{\textbf{Figure 5:} This figure best summarizes the results, and a version of this should be the Graphical Abstract, with legends for the color cods and spelling out the missing data parameters, and the Y axis could just be 'clade similarity index' or 'rouge taxon index' so that the reader doesn't need to know what the names of the metrics are to grasp the graphic. But concerning the Normalized Triplets metric, according to the text and appendix, if this index is 0, the similarity of the missing data tree to the best tree is no more than random, and < 0 it is more different than random. For the Bayesian analyses, many results had values < 0 but the axis is set to 0. Why not illustrate the distribution of values that is < 0? From this result, it appears the Bayesian analysis does quite poorly with missing data in terms of rouge taxa, and worse than max. likelihood. This figure also suggests to me that the methods perform pretty badly with missing data, in contrast to previous findings that missing data are not overly deleterious. For example, when all data are complete except Mf=50\%, the normalized RF metric is <0.8. Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005), although for small datasets such as the one used in this study, the results show a similar decline in accuracy. This is a point that should be expanded upon in the discussion - how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?}}

We thank the reviewer for this excellent suggestion and modified our graphical abstract.
%To do (writing)
%Missing data for up to 50\% of characters has been suggested to still have high accuracy in uncovering the true tree (Wiens 2003; Wiens 2005)
%how these results compare to other missing data studies and what do they tell us about the performance of the inference methods with missing data?

\item{\textcolor{blue}{\textbf{Figure 6:} I think this may be the most important figure because it is demonstrating the probability of overlap between the distributions of the best trees and the missing data trees. It took me a long time to figure it out, however, and to figure out why the top of the triangle is all orange. I would have expected something more like in Appendix C Figure C2 or Fig. C3. At first I understood this to indicate high probability of overlap of the missing data trees which had the most missing data with the best tree. But after re-reading the legend and methods several times, I am interpreting it as follows: the top of the triangle indicates a similarly poor fit of the distributions of missing data trees to the best tree under all parameter combinations once there is a high proportion of missing data. The point is raised in the Results that once missing data in the living portion is > 50\% then they all perform poorly, but it was not easy to grasp. Somehow this figure needs to be spelled out better in the caption and results so the reader can grasp it more quickly and easily without having to go back to the methods section to completely figure out what is being represented.}}

%To do (writing)

@@

%To do (writing)

\item{\textcolor{blue}{\textbf{Table 1:} The wording of the caption is off - the second paragraph is awkward. 
Are these values the probabilities of overlap between the distributions of the 'best' tree versus trees from each inference method pooled across all combinations of missing data? That is what is described in methods line 397 but it doesn't say so in the caption. Two things:\\
1)  Comparing to the tables in Appendix C, the distribution of probabilities of overlap between max. likelihood and Bayesian trees are higher in most missing data schemes in tables in the Appendices than the overall values given in Table 1. How could the max. likelihood and Bayesian trees be so different?\\
2)  Pooling across all missing data schemes seems to wash out the effect of tree-inference method. I am not sure what a better way to present these would be, but perhaps some kind of matrix where the cells are the joint probabilities of overlap of each method with each missing data category, and then the marginal probabilities would be the probabilities of overlap for each method, averaged over missing data categories, as well as the marginal probabilities of overlap for each missing data category, averaged across methods.  e.g, RF distance:\\
Missing data:   0   25  50  75  Marginal \\
BI  1   0.9 0.8 0.75    0.863 \\
ML  1   0.9 0.75    0.25    0.725 \\
Marginal    1   0.9 0.78    0.5  \\
This may not be feasible in the text due to space limitations, maybe tables like these could be added to an appendix.}}

\item{\textcolor{blue}{\textbf{Appendix C:} Tables 1-3 repeats 'of the'. Should label the first two rows as 'Best maximum likelihood tree' and 'Best Bayesian inference tree'}}

We removed the repetition in the caption. %Not sure I understand the second part of the comment.

\item{\textcolor{blue}{\textbf{Appendix D:} I find the online code repository thorough, but the repository is loaded with many files, especially for graphing. If it is possible to condense these files it may be easier for a reader to duplicate the study.}}

### 2015-06-30 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixC.tex
Del with all the minor changes

%Summary of metrics values for all parameters combinations
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
\label{Tab_Supp_summary_metric_allparam}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
                                        & $Tr$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
                                        & $Tr$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
                                        & $Tr$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
                                        & $Tr$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the ML parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
                                        & $Tr$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
  Bayesian consensus                    & $RF$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
                                        & $Tr$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
                                        & $Tr$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
                                        & $Tr$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MF parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MF}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
                                        & $Tr$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
                                        & $Tr$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
                                        & $Tr$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MC parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MC}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
                                        & $Tr$ & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
                                        & $Tr$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
                                        & $Tr$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

>>>>

%Summary of metrics values for all parameters combinations
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr).}
\label{Tab_Supp_summary_metric_allparam}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
                                        & $Tr$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
                                        & $Tr$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
                                        & $Tr$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
                                        & $Tr$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the ML parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{L}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
                                        & $Tr$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
  Bayesian consensus                    & $RF$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
                                        & $Tr$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
                                        & $Tr$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
                                        & $Tr$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MF parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{F}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MF}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
                                        & $Tr$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
                                        & $Tr$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
                                        & $Tr$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

%Summary of metrics values for the MC parameter
\begin{landscape}
\begin{table}[ht]
\caption{Summary of the comparisons between the "best" tree and the "missing data" trees for each different tree inference method using either the Normalised Robinson-Foulds metric (RF) or the Normalised Triplets metric (Tr) for the $M_{C}$ missing data parameter only.}
\label{Tab_Supp_summary_metric_MC}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
                                        & $Tr$ & 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
                                        & $Tr$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
  Maximum Likelihood bootstraps         & $RF$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
                                        & $Tr$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
                                        & $Tr$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
\end{table}
\end{landscape}

### 2015-06-30 Manuscript/MPESubmission/Revision/TEM_revised_manuscript.tex
Del with all the minor changes

\section{Conclusions}
Missing data in Total Evidence matrices is not a problem for recovering the ``best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
Therefore we advise filling as many gaps in Total Evidence matrices as possible.
Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible.
Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for furhter analyses.
The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.

@@

\noindent
\textbf{Figure 2:} Bhattacharyya Coefficient calculation outline 1. A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B. Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.\\

\noindent
\textbf{Figure 3:} Bhattacharyya Coefficient calculation outline 2. A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g. Maximum Likelihood or Bayesian). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$. Note that this is similar to performing a two-sided t-test, but we use the Bhattacharyya Coefficient here because we are comparing distributions not means.\\

\bigskip
\noindent
\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue). The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.

>>>>

\section{Conclusions}
Missing data in Total Evidence matrices is not a problem for recovering the ``best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
Therefore we advise filling as many gaps in Total Evidence matrices as possible.
Because it can be difficult to do for fossil taxa, we recommend coding as many morphological characters as possible for living taxa.
Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the ``best" tree topology than the Maximum Likelihood trees.
Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed for further analyses.
The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data.
Hopefully, using these approaches will greatly improve our understanding of macroevolutionary patterns and processes.

@@

\noindent
\textbf{Figure 2:} Bhattacharyya Coefficient calculation outline 1. A, B and C are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two combinations of missing data parameters, for example, BC(A,B) is the probability of overlap between the distributions A and B.\\

\noindent
\textbf{Figure 3:} Bhattacharyya Coefficient calculation outline 2. A and B are distributions of tree similarity metrics (Normalised Robinson-Foulds or Normalised Triplets metrics) for any combination of missing data parameters (e.g. $M_{L}$ = 10\%, $M_{F}$ = 50\%, $M_{C}$ = 75\%). \textbf{(x)} and \textbf{(y)} are two different tree inference methods (e.g. Maximum Likelihood or Bayesian). The Bhattacharyya Coefficient (BC) is the overlap of the distribution of tree similarity metrics between two methods for the same combination of missing data parameters, for example, BC($A_{x}$,$A_{y}$) is the probability of overlap of the distribution A for methods $x$ and $y$.\\

\bigskip
\noindent
\textbf{Figure 4:} The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (red), Maximum Likelihood bootstrap trees (blue) and Bayesian posterior tree distributions (orange). The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds metric (upper row) and Normalised Triplets metric (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.
